Spatial Analysis

### 2021-10-05 Scripts/spatial_analisis.Rmd
interpolation figure (i.e. figure 2)

# Load all sp in one table
historic_spp <- bind_rows(lapply(taxon_sw, fread)) %>% 
  left_join(grids,
            by = c("index","lon","lat")
            ) %>%
  filter(!is.na(spatial))

@@

unique(historic_spp$spp)
unique(historic_spp$region)
unique(historic_spp$region)

@@

# Interpolation data
interpol_result <- historic_spp %>% 
  mutate(
    label = ifelse(year >= 1980 & year <= 2001,"Reference",
                   ifelse(year > 2001,"Today",NA)
    )
  ) %>% 
  filter(!is.na(label),
         region == "Northeast US Spring") %>% 
  group_by(label,index,spp,lat,lon) %>% 
  summarise(mean_cpue = mean(value,na.rm=T)) %>% 
  filter(!is.na(mean_cpue))

@@

# Estimate total interpolation
total_inter <- historic_spp %>% 
  mutate(
    label = ifelse(year >= 1980 & year <= 2001,"Reference",
                   ifelse(year > 2001,"Today",NA)
    )
  ) %>% 
  filter(!is.na(label),
         region == "Northeast US Spring") %>% 
  group_by(spp,index) %>% 
  summarise(mean_cpue = mean(value,na.rm=T)) %>% 
  group_by(spp) %>% 
  summarise(total_cpue = sum(mean_cpue, na.rm = T))

# Estimate centroid based on top 10 grids with higer abundance
centroid_db <- historic_spp %>% 
  mutate(
    label = ifelse(year >= 1980 & year <= 2001,"Reference",
                   ifelse(year > 2001,"Today",NA)
    )
  ) %>% 
  filter(!is.na(label),
         region == "Northeast US Spring") %>% 
  group_by(label,index,spp,lat,lon) %>% 
  summarise(mean_cpue = mean(value,na.rm=T)) %>% 
  filter(!is.na(mean_cpue)) %>% 
  left_join(total_inter) %>% 
  mutate(
    per_cpue = round((mean_cpue/total_cpue)*100,2)
  ) %>% 
  group_by(spp,label) %>% 
  top_n(10,per_cpue) %>% 
  group_by(spp,label) %>% 
  summarise(
    lat = mean(lat),
    lon = mean(lon)
  )

inter_map <- ggplot() +
   geom_tile(data = subset(interpol_result, mean_cpue == 0),
            aes(
              x = lon,
              y = lat
            ),
            fill = "grey90",
              color = "grey90"
  ) + 
  geom_tile(data = subset(interpol_result, mean_cpue > 0),
            aes(
              x = lon,
              y = lat,
              fill = mean_cpue,
              color = mean_cpue
            )
  ) + 
  geom_point(data = centroid_db,
            aes(
              x = lon,
              y = lat
            ),
            shape = 3,
            color = "white"
  ) + 
  geom_sf(data = land_sf,aes()) +
  labs(x = "Longitude",
         y = "Latitude") +
  # ggsflabel::geom_sf_label_repel(data = land_sf, aes(label= abrev), 
  #                                  size = 4,
  #                                  box.padding = 0.10,
  #                                  hjust = 1) +
  scale_fill_viridis("Mean\nInterpolation",
                     limits = c(0,40)) +
  scale_color_viridis("Mean\nInterpolation",
                      limits = c(0,40)) +
  facet_grid(spp~label) +
  my_ggtheme_m(map_type = "na", leg_pos = "right",ax_tx_s = 7,ax_tl_s = 10,leg_tx_s = 10) #+
  # theme(legend.key.width = unit(3,"line"))

ggsave("interpolation_fig.jpg",
       inter_map,
       path = my_path("R","Figures"),
       width = 6,
       height = 6)
```

>>>>

# Load all sp in one table
historic_spp <- bind_rows(lapply(taxon_sw, fread)) %>% 
  left_join(grids,
            by = c("index","lon","lat")
            ) %>%
  filter(!is.na(spatial)) %>% 
  mutate(
    season = ifelse(str_detect(region,"Spring"),"Spring","Fall")
  )

@@

unique(historic_spp$spp)
unique(historic_spp$region)
unique(historic_spp$season)

@@

# Interpolation data
interpol_result <- historic_spp %>% 
  mutate(
    label = ifelse(year >= 1980 & year <= 2001,"Reference",
                   ifelse(year > 2001,"Today",NA)
    )
  ) %>% 
  filter(!is.na(label)) %>% 
  group_by(label,index,spp,season,lat,lon) %>% 
  summarise(mean_cpue = mean(value,na.rm=T)) %>% 
  filter(!is.na(mean_cpue))

@@

# Estimate total interpolation
total_inter <- historic_spp %>% 
  mutate(
    label = ifelse(year >= 1980 & year <= 2001,"Reference",
                   ifelse(year > 2001,"Today",NA)
    )
  ) %>% 
  filter(!is.na(label)) %>% 
  group_by(spp,index,season) %>% 
  summarise(mean_cpue = mean(value,na.rm=T)) %>% 
  group_by(spp,season) %>% 
  summarise(total_cpue = sum(mean_cpue, na.rm = T))

# Estimate centroid based on top 10 grids with higer abundance
centroid_db <- historic_spp %>% 
  mutate(
    label = ifelse(year >= 1980 & year <= 2001,"Reference",
                   ifelse(year > 2001,"Today",NA)
    )
  ) %>% 
  filter(!is.na(label)) %>% 
  group_by(label,index,spp,season,lat,lon) %>% 
  summarise(mean_cpue = mean(value,na.rm=T)) %>% 
  filter(!is.na(mean_cpue)) %>% 
  left_join(total_inter) %>% 
  mutate(
    per_cpue = round((mean_cpue/total_cpue)*100,2)
  ) %>% 
  group_by(spp,label,season) %>% 
  top_n(10,per_cpue) %>% 
  group_by(spp,label,season) %>% 
  summarise(
    lat = mean(lat),
    lon = mean(lon)
  )

pal <- wes_palette("Zissou1",100,type = "continuous")

inter_map <- ggplot() +
  geom_tile(data = subset(interpol_result, mean_cpue == 0),
            aes(
              x = lon,
              y = lat
            ),
            fill = "grey90",
            color = "grey90"
  ) + 
  geom_tile(data = subset(interpol_result, mean_cpue > 0),
            aes(
              x = lon,
              y = lat,
              # fill = log10(mean_cpue),
              # color = log10(mean_cpue)
              fill = mean_cpue,
              color = mean_cpue
            )
  ) + 
  geom_point(data = centroid_db,
             aes(
               x = lon,
               y = lat
             ),
             shape = 3,
             color = "white"
  ) + 
  geom_sf(data = land_sf,aes()) +
  labs(x = "Longitude",
       y = "Latitude") +
  # ggsflabel::geom_sf_label_repel(data = land_sf, aes(label= abrev), 
  #                                  size = 4,
  #                                  box.padding = 0.10,
  #                                  hjust = 1) +
  # Viridis option
scale_fill_viridis("CPUE",
                   limits = c(0,40),
                   # limits = c(-5,1), # for log10
                   direction = -1,
                    end = 0.9
                   ) +
scale_color_viridis("CPUE",
                    limits = c(0,40),
                    # limits = c(-5,1), # for log10
                    direction = -1,
                    end = 0.9
                    ) +
# wes anderson option
# scale_fill_gradientn("CPUE",
#                      colours = pal,
#                      limits = c(0,40) # For cpue
#                      # limits = c(-5,1) # for log10
#                      
# ) +
#   scale_color_gradientn("CPUE",
#                         colours = pal,
#                         limits = c(0,40) # for cpue
#                         # limits = c(-5,1) # for log10
#   ) +
  facet_grid(spp~season+label) +
  my_ggtheme_m(map_type = "na", leg_pos = "right",ax_tx_s = 7,ax_tl_s = 10,leg_tx_s = 12) #+
# theme(legend.key.width = unit(3,"line"))

ggsave("interpolation_vi.jpg",
       inter_map,
       path = my_path("R","Figures"),
       width = 9,
       height = 7)
```
